An introduction to Docker for reproducible research

### 2014-11-06 docker-paper/manuscript.Rmd
further edits

Consider the following Dockerfile, which

```
FROM rocker/ropensci:latest
RUN git clone https://github.com/ropensci/RNeXML.git /rnexml
RUN Rscript -e 'devtools::install("/rnexml"); devtools::check("/rnexml")'
```

A Dockerfile can have only a single `FROM` line, while sometimes it may
be necessary to build on the computational environment provided by more
than one container. To address this, Docker defines a syntax for linking
multiple containers together:

Each one acts as a building block providing just what is necessary to
run one particular service or element, and exposing just what is necessary
to link it together with other blocks.

```bash
docker run --name tex -v /usr/local/texlive rocker/texlive
docker run -dP --volumes-from tex \
-e PATH=$PATH:/usr/local/texlive/2014/bin/x86_64-linux/ \
rocker/rstudio 
```

For instance, one could have one
container running a PostgreSQL database which serves data to another container
running a python environment to analyze the data:

>>>>

<!--
Consider the following Dockerfile to illustrate this modular reuse. This
is the complete Dockerfile required to create the computational environment
required for a particular manuscript:

```bash
FROM rocker/ropensci:latest
MAINTAINER Carl Boettiger cboettig@ropensci.org
RUN git clone https://github.com/ropensci/RNeXML.git /rnexml
RUN Rscript -e 'devtools::install("/rnexml")'
```

The first line contains the FROM directive, indicating that this Dockerfile
builds on the existing Docker image `rocker/ropensci`, hosted on the Docker
Hub.  For this approach to be effective for reproducible one must take care 
that the base image and it's Dockerfile recipe are available from a persistant
archive.

The second line is optional, declaring minimal metadata regarding the 
contact information for the image maintainer. The third line copys the
project onto the image and the fourth installs the necessary software that
has not been provided by the base image. Projects using less standard
environments may require more instructions.  
-->

While a Dockerfile can have only a single `FROM` line, sometimes it may
be necessary to build on the computational environment provided by more
than one container. To address this, Docker defines a syntax for linking
multiple containers together.  This allows each container to act as a
building block providing just what is necessary to run one particular
service or element, and exposing just what is needed to link it together
with other blocks.  For instance, one could have one container running
a PostgreSQL database which serves data to another container running a
python environment to analyze the data:

<!--
```bash
docker run --name texlive -v /usr/local/texlive rocker/texlive
docker run -dP --volumes-from texlive \
-e PATH=$PATH:/usr/local/texlive/2014/bin/x86_64-linux/ \
rocker/rstudio 
```
-->
